The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-04-18 src/ms.tex
JWST

\paragraph{A High-Level Interface to World Coordinate Systems}
Astronomical data are often provided alongside information about the
correspondence between ``real-world'' and pixel coordinates. This mapping is the
essence of the World Coordinate System (WCS) concept. From its inception,
the \astropysubpkg{wcs} sub-package allowed access to WCS information provided
in, e.g., FITS files, but it became clear that other WCS standards and
representations had to be supported for new missions and observatories (e.g.,
the James Webb Space Telescope and the Rubin Observatory). To harmonize these
needs, a new high-level interface was created based on a formal design first
proposed in an \astropy\ Proposal for Enhancement (\astropyAPE{14}). The hope is
that by having a formal design, other packages implementing WCS objects can
straightforwardly modify their classes to conform to the new interface or build
thin wrappers that conform. The implementation in \astropysubpkg{wcs} interacts
well with other \astropypkg\ objects such as \astropySkyCoord\ and \astropyTime.
Future work includes better integration of this interface with the rest of
the Project.

@@

Several of the coordinated and affiliated packages described in
\cite{astropy:2018} have had substantial improvements. \texttt{astroquery}
\citep{astroquery} has added access to roughly a dozen new missions and data
services, including the James Webb Space Telescope (JWST) archive, and the
project has switched to a continuous release model: every time a change is
committed to the main development branch it is published on the Python Package
Index (PyPI) and available for installation. (However, formal releases are still
done a few times per year.) \texttt{photutils} \citep{photutils} released its
first stable version, indicating that the API will change less frequently, and
there have been several significant performance improvements. \texttt{ccdproc}
\citep{ccdproc}  released a new major version, bringing better performance to
some image combination operations. The \texttt{regions} package
\citep{pyregions}, for manipulating ds9-, CASA-, or FITS-style region definitions \citep{ds9},
added new ways to manipulate regions, introduced new region types, and obtained
expanded visualization and interactive region editing capabilities.
\texttt{reproject}'s \citep{reproject} major new feature is a function to align
and co-add images to create a mosaic; better support for parallelization was
also added. \texttt{specutils}, the package that defines containers for 1D and
2D spectra \citep{specutils},  had its first stable release and the addition
of classes to read JWST data. \texttt{stingray} has had a major performance
overhaul. The package \texttt{gammapy} \citep{gammapy} has had major performance
improvements and has unified its API in preparation for its first stable
release.

@@

\subsubsection{James Webb Space Telescope}
% \secauthor{Larry Bradley}

The James Webb Space Telescope (JWST) is a 6.5-meter space-based
infrared telescope that will provide unprecedented resolution and
sensitivity from 0.6--28 microns. JWST will enable a broad range
of scientific investigations from exoplanets and their atmospheres
to the formation of galaxies in the very early universe. Its four
key scientific goals are to study the first light from stars and
galaxies, the assembly and evolution of galaxies, the birth of stars and
protoplanetary systems, and planetary systems and the origins of life.

>>>>

\paragraph{A High-Level Interface to World Coordinate Systems}
Astronomical data are often provided alongside information about the
correspondence between ``real-world'' and pixel coordinates. This mapping is the
essence of the World Coordinate System (WCS) concept. From its inception,
the \astropysubpkg{wcs} sub-package allowed access to WCS information provided
in, e.g., FITS files, but it became clear that other WCS standards and
representations had to be supported for new missions and observatories (e.g.,
JWST and the Rubin Observatory). To harmonize these
needs, a new high-level interface was created based on a formal design first
proposed in an \astropy\ Proposal for Enhancement (\astropyAPE{14}). The hope is
that by having a formal design, other packages implementing WCS objects can
straightforwardly modify their classes to conform to the new interface or build
thin wrappers that conform. The implementation in \astropysubpkg{wcs} interacts
well with other \astropypkg\ objects such as \astropySkyCoord\ and \astropyTime.
Future work includes better integration of this interface with the rest of
the Project.

@@

Several of the coordinated and affiliated packages described in
\cite{astropy:2018} have had substantial improvements. \texttt{astroquery}
\citep{astroquery} has added access to roughly a dozen new missions and data
services, including the JWST archive, and the
project has switched to a continuous release model: every time a change is
committed to the main development branch it is published on the Python Package
Index (PyPI) and available for installation. (However, formal releases are still
done a few times per year.) \texttt{photutils} \citep{photutils} released its
first stable version, indicating that the API will change less frequently, and
there have been several significant performance improvements. \texttt{ccdproc}
\citep{ccdproc}  released a new major version, bringing better performance to
some image combination operations. The \texttt{regions} package
\citep{pyregions}, for manipulating ds9-, CASA-, or FITS-style region definitions \citep{ds9},
added new ways to manipulate regions, introduced new region types, and obtained
expanded visualization and interactive region editing capabilities.
\texttt{reproject}'s \citep{reproject} major new feature is a function to align
and co-add images to create a mosaic; better support for parallelization was
also added. \texttt{specutils}, the package that defines containers for 1D and
2D spectra \citep{specutils},  had its first stable release and the addition
of classes to read JWST data. \texttt{stingray} has had a major performance
overhaul. The package \texttt{gammapy} \citep{gammapy} has had major performance
improvements and has unified its API in preparation for its first stable
release.

@@

\subsubsection{JWST}
% \secauthor{Larry Bradley}

JWST is a 6.5-meter space-based
infrared telescope that will provide unprecedented resolution and
sensitivity from 0.6--28 microns. JWST will enable a broad range
of scientific investigations from exoplanets and their atmospheres
to the formation of galaxies in the very early universe. Its four
key scientific goals are to study the first light from stars and
galaxies, the assembly and evolution of galaxies, the birth of stars and
protoplanetary systems, and planetary systems and the origins of life.
